forrechelle

### 2019-12-20 analysis/06_DESeq2_ALL.Rmd
now with some corerlations

char <- plot_grid(
          b + theme(axis.title.x = element_blank(), axis.text.x = element_blank()),
          c + theme(axis.title = element_blank(), axis.text = element_blank()) ,
          d + theme(axis.title = element_blank(), axis.text = element_blank()), 
          f + theme(axis.title = element_blank(), axis.text = element_blank()),
          g + theme(axis.title = element_blank(), axis.text = element_blank()),
          h + theme(axis.title = element_blank(), axis.text = element_blank()),
          i + theme(axis.title = element_blank(), axis.text = element_blank()),
          nrow = 1, rel_widths = c(1.2,1,1,1,1,1,1))

remove <- plot_grid( 
          a + theme(axis.title.x = element_blank(), axis.text.x = element_blank()),
          NULL, 
          j + theme(axis.title = element_blank(), axis.text = element_blank()) , 
          k + theme(axis.title = element_blank(), axis.text = element_blank()) , 
          l + theme(axis.title = element_blank(), axis.text = element_blank()) , 
          m + theme(axis.title = element_blank(), axis.text = element_blank()), 
          NULL, nrow = 1, rel_widths = c(1.2, 1, 1,1,1,1,1))

timing <- plot_grid( 
          n , 
          o + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) , 
          NULL,
          p + theme(axis.title.y = element_blank(), axis.text.y = element_blank()), 
          q + theme(axis.title.y = element_blank(), axis.text.y = element_blank())+
            labs(x = "Log fold change, pituitary") , 
           NULL,
          r + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) , 
          s + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
            nrow = 1, rel_widths = c(1.2,1, 0.5, 1,1,0.5, 1,1))

>>>>

char <- plot_grid(
          b + theme(axis.title.x = element_blank(), axis.text.x = element_blank()),
          c + theme(axis.title = element_blank(), axis.text = element_blank()) ,
          d + theme(axis.title = element_blank(), axis.text = element_blank()), 
          f + theme(axis.title = element_blank(), axis.text = element_blank()),
          g + theme(axis.title = element_blank(), axis.text = element_blank()),
          h + theme(axis.title = element_blank(), axis.text = element_blank()),
          i + theme(axis.title = element_blank(), axis.text = element_blank()),
          nrow = 1, rel_widths = c(1.3,1,1,1,1,1,1))

remove <- plot_grid( 
          a + theme(axis.title.x = element_blank(), axis.text.x = element_blank()),
          NULL, 
          j + theme(axis.title = element_blank(), axis.text = element_blank()) , 
          k + theme(axis.title = element_blank(), axis.text = element_blank()) , 
          l + theme(axis.title = element_blank(), axis.text = element_blank()) , 
          m + theme(axis.title = element_blank(), axis.text = element_blank()), 
          NULL, nrow = 1, rel_widths = c(1.3, 1, 1,1,1,1,1))

timing <- plot_grid( 
          n , 
          o + theme(axis.title.y = element_blank(), axis.text.y = element_blank())+
            labs(x = "Log fold change,")   , 
          NULL,
          p + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) +
            labs(x = "pituitary") , 
          q + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) ,
           NULL,
          r + theme(axis.title.y = element_blank(), axis.text.y = element_blank()) , 
          s + theme(axis.title.y = element_blank(), axis.text.y = element_blank()),
            nrow = 1, rel_widths = c(1.3,1, 0.5, 1, 1, 0.5, 1, 1))

### 2019-12-20 analysis/20_forrechelle.Rmd
now with some corerlations

```{r setup, include=FALSE}
library(tidyverse)
library(ggimage)
library(cowplot)

source("../R/themes.R")

@@

candidategenes <- c( "NR3C1", "NR3C2", "PRLR", "PRL")
timepoints <- c("lay" , "inc.d9", "hatch", "n9")
hormones <- c("PRL", "CORT")

@@

df <- df %>% filter(tissue != "gon")
head(df)

df2 <-  df  %>%
  filter(X1 %in% candidategenes) %>%
  pivot_longer(cols = L.Blu13_male_gonad_control.NYNO:y98.o50.x_male_pituitary_inc.d3, 
               names_to = "sample", values_to = "vsd") %>%
   mutate(sex = sapply(strsplit(as.character(sample),'_'), "[", 2)) %>%
   mutate(treatment = sapply(strsplit(as.character(sample),'_'), "[", 4))  %>%
   mutate(treatment = sapply(strsplit(as.character(treatment),'.NYNO'), "[", 1)) %>%
  mutate(bird = sapply(strsplit(as.character(sample),'_'), "[", 1)) %>%
  filter(treatment %in% timepoints) %>%
  select(bird, sex, treatment, tissue, X1, vsd) %>%
  rename(gene = X1) %>%
  drop_na()
head(df2)

df3 <- df2 %>% 
  mutate(tissue = fct_relevel(tissue,  "hyp", "pit", "gon")) %>% 
  mutate(treatment = fct_relevel(treatment, timepoints)) %>% 
  group_by(sex, treatment, tissue, gene)  %>% 
  summarize(m = mean(vsd, na.rm = T), se = sd(vsd,  na.rm = T)/sqrt(length(vsd))) %>%
  mutate(image = "../figures/images/DoveParentsRNAseq_note.png")
head(df3)

@@

p <-  df3 %>%
  filter(tissue == i) %>%
  ggplot(aes(x = treatment, y = m)) +
  geom_image(aes(image=image, color = gene), size=0.1)  +

  scale_alpha_manual(values = c(0.5, 1)) +
  labs(subtitle = i, y = "average expression", x = "parental stage")

pimage <- axis_canvas(p, axis = 'x') + 
  draw_image("../figures/images/icons/DoveParentsRNAseq_lay.png", x = 0.5, scale = 1 ) +
  draw_image("../figures/images/icons/DoveParentsRNAseq_incubation.png", x = 1.5, scale = 1) +
  draw_image("../figures/images/icons/DoveParentsRNAseq_hatch.png", x = 2.5, scale = 1) +
  draw_image("../figures/images/icons/DoveParentsRNAseq_chicklings.png", x = 3.5, scale = 1)

p2 <- ggdraw(insert_xaxis_grob(p, pimage, position = "bottom"))
plot(p2)

}

@@

  p <-  df3 %>%
    filter(tissue == i) %>%
    ggplot(aes(x = treatment, y = m)) +
    geom_errorbar(aes(ymin=m-se, ymax=m+se, color=gene), width=.1) +
    geom_point(size = 1, aes(color = gene)) +
    geom_line(aes(x = as.numeric(treatment), y = m, color = gene)) +
    scale_alpha_manual(values = c(0.5, 1)) +
    labs(subtitle = i, y = "average expression", x = "parental stage") +
    facet_wrap(~sex, nrow = 1) +
    theme_B3()
  plot(p)

}

>>>>

```{r setup, include=FALSE}
library(tidyverse)
library(ggimage)
library(cowplot)
library(forcats)
library(corrr)

source("../R/themes.R")
source("../R/functions.R")

@@

candidategenes <- c( "NR3C1", "NR3C2", "PRLR", "PRL", "POMC" ,"MC4R", "VIP", "BRCA1")
timepoints <- c("lay" , "inc.d9", "hatch", "n9")
hormones <- c("PRL", "CORT")

@@

genes <- as.data.frame(df$X1)

df2 <-  df  %>%
  filter(X1 %in% candidategenes) %>%
  pivot_longer(cols = L.Blu13_male_gonad_control.NYNO:y98.o50.x_male_pituitary_inc.d3, 
               names_to = "sample", values_to = "vsd") %>%
   mutate(sex = sapply(strsplit(as.character(sample),'_'), "[", 2)) %>%
   mutate(treatment = sapply(strsplit(as.character(sample),'_'), "[", 4))  %>%
   mutate(treatment = sapply(strsplit(as.character(treatment),'.NYNO'), "[", 1)) %>%
  mutate(bird = sapply(strsplit(as.character(sample),'_'), "[", 1)) %>%
  filter(treatment %in% timepoints) %>%
  select(bird, sex, treatment, tissue, X1, vsd) %>%
  mutate(tissue = fct_recode(tissue, "hypothalamus" = "hyp",
                    "pituitary" = "pit",
                    "gonads" = "gon"
                    )) %>%
  rename(gene = X1) %>%
  drop_na() %>%
  droplevels()
head(df2)

df2$treatment <- factor(df2$treatment, levels = alllevels)

df3 <- df2 %>% 
  mutate(treatment = fct_relevel(treatment, timepoints)) %>% 
  group_by(sex, treatment, tissue, gene)  %>% 
  summarize(m = mean(vsd, na.rm = T), se = sd(vsd,  na.rm = T)/sqrt(length(vsd))) %>%
  mutate(image = "../figures/images/DoveParentsRNAseq_note.png")
head(df3)

@@

  p <-  df3 %>%
    filter(tissue == i) %>%
    ggplot(aes(x = treatment, y = m)) +
    geom_errorbar(aes(ymin=m-se, ymax=m+se, color=gene), width=.1) +
    geom_point(size = 1, aes(color = gene)) +
    geom_line(aes(x = as.numeric(treatment), y = m, color = gene)) +
    scale_alpha_manual(values = c(0.5, 1)) +
    labs(subtitle = i, y = "average expression", x = "parental stage") +
    facet_wrap(~sex, nrow = 1) +
    theme_B3()
 print(p)
}

@@

  p <-  df2 %>%
    filter(tissue == i) %>%
    ggplot(aes(x = treatment, y = vsd, fill = gene, color = sex)) +
    geom_boxplot() +
    scale_alpha_manual(values = c(0.5, 1)) +
    labs(subtitle = i, y = "expression", x = "parental stage") +
    facet_wrap(~gene, scales = "free_y") +
    theme_B3() +
    scale_color_manual(values = sexcolors)
 print(p)
}

dft <- df %>% filter(X1 %in% candidategenes) %>%
  unite("tissuegene", tissue:X1, remove = FALSE) %>%
  select(-tissue, -`X1`, -filename)  %>%
  pivot_longer(-tissuegene, names_to = "samples", values_to = "vsd") %>%
  drop_na() %>%
  mutate("bird" = sapply(strsplit(as.character(samples),'\\_'), "[", 1)) %>%
  select(bird,tissuegene,vsd) %>%
  pivot_wider(id_cols = bird, names_from = tissuegene, values_from = vsd)

dft <- as.data.frame(dft)
row.names(dft) <- dft$bird
dft$bird <- NULL

x <- correlate(dft)

x %>% network_plot.cor_df(min_cor = .2)

x %>%
  focus(pit_PRL) %>%
  mutate(rowname = reorder(rowname, pit_PRL)) %>%
  ggplot(aes(rowname, pit_PRL)) +
    geom_col() + coord_flip() +
  labs(y = "corrleation with pituitiary PRL", x = "candidate genes by tissue") +
  theme_B3() + 
  theme(legend.position = "none")
```
